---
title: "Demo page"
format: 
  live-html:
    theme: cosmo
    css: assets/webex_2.css
    include-after-body: assets/webex_2.js
engine: knitr
---

{{< include ./_extensions/r-wasm/live/_knitr.qmd >}}


## Exercise interactive code with hints and solution

:::: {.panel-tabset}

## Exercise

```{webr}
#| setup: true
#| exercise: ex_4
library(idarps)
```

```{webr}s
#| exercise: ex_4
data("pharmacy")
# Construct attendance by day
monday = na.omit(pharmacy$attendance[pharmacy$weekday == "Monday"])
tuesday = na.omit(pharmacy$attendance[pharmacy$weekday == "Tuesday"])
wednesday = na.omit(pharmacy$attendance[pharmacy$weekday == "Wednesday"])
thursday = na.omit(pharmacy$attendance[pharmacy$weekday == "Thursday"])
friday = na.omit(pharmacy$attendance[pharmacy$weekday == "Friday"])
saturday = na.omit(pharmacy$attendance[pharmacy$weekday == "Saturday"])
sunday = na.omit(pharmacy$attendance[pharmacy$weekday == "Sunday"])
idarps::___________(monday, tuesday, wednesday,
                    thursday, friday, saturday, sunday)
```

## Hints

::: { .hint exercise="ex_4"}
Consider using the `boxplot_w_points()` function from package `idarps`.

```r
?idarps::boxplot_w_points
```
:::

## Solution

::: { .solution exercise="ex_4" }
Use the function `boxplot_w_points()` from package `idarps`:

```r
data("pharmacy")

# Construct attendance by day
monday = na.omit(pharmacy$attendance[pharmacy$weekday == "Monday"])
tuesday = na.omit(pharmacy$attendance[pharmacy$weekday == "Tuesday"])
wednesday = na.omit(pharmacy$attendance[pharmacy$weekday == "Wednesday"])
thursday = na.omit(pharmacy$attendance[pharmacy$weekday == "Thursday"])
friday = na.omit(pharmacy$attendance[pharmacy$weekday == "Friday"])
saturday = na.omit(pharmacy$attendance[pharmacy$weekday == "Saturday"])
sunday = na.omit(pharmacy$attendance[pharmacy$weekday == "Sunday"])
idarps::boxplot_w_points(monday, tuesday, wednesday, #<1>
                         thursday, friday, saturday, sunday)           
                     
          
```
1. Use function `boxplot_w_points()` from the `idarps` package to create a boxplot with points overlaid.

:::

::::

## Python code chunk
```{python}
1+1
```
## R code chunk
```{r}
1+1
```





## Iframe 

<iframe src="https://raw.githack.com/stephaneguerrier/data_analytics/master/Lecture1.html#1" width="710" height="530">
</iframe>

## Local pdf slides


<embed src="cours_3.pdf" type="application/pdf" width="100%" height="600px" />


## Python interactive

```{pyodide}
for x in range(5):
  print(x ** 2)
import matplotlib.pyplot as plt
plt.plot([1, 2, 3, 4], [1, 4, 9, 16])
plt.ylabel('Magic Numbers')
plt.show()
```


## R interactive
```{webr}
library(idarps)
data("pharmacy")

# Construct attendance by day
monday = na.omit(pharmacy$attendance[pharmacy$weekday == "Monday"])
tuesday = na.omit(pharmacy$attendance[pharmacy$weekday == "Tuesday"])
wednesday = na.omit(pharmacy$attendance[pharmacy$weekday == "Wednesday"])
thursday = na.omit(pharmacy$attendance[pharmacy$weekday == "Thursday"])
friday = na.omit(pharmacy$attendance[pharmacy$weekday == "Friday"])
saturday = na.omit(pharmacy$attendance[pharmacy$weekday == "Saturday"])
sunday = na.omit(pharmacy$attendance[pharmacy$weekday == "Sunday"])
idarps::boxplot_w_points(monday, tuesday, wednesday, thursday, friday, saturday, sunday)
```





## MCQ 
```{r, echo = FALSE, results='asis'}
knitr::opts_chunk$set(echo = FALSE)
source("assets/q_a_env.R")
```

<!-- MathJax for Rendering Math -->
<script type="text/javascript" async
  src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/3.2.2/es5/tex-mml-chtml.js">
</script>


## Question 1

Here is a question.

```{r, results='asis'}
# Define questions choice and individual feedback for each question
opts <- list(
  list(value = "\\(\\frac{1}{10}\\)",
       feedback = "Too small ! ðŸ˜±"),
  list(value = "\\(\\frac{3}{10}\\)",
       feedback = "Close! "),
  answer = list(
    value = "\\(\\frac{2}{5}\\)",
    feedback = "correct ! ðŸ‰ \\(\\frac{52}{100}\\)"
  ),
  list(
    value = "\\(\\frac{52}{100}\\)",
    feedback = "Not correct with \\(\\frac{52}{100}\\)! ðŸ¤“"
  )
)
# render QA environment
render_q_a_env(opts)
small_v_space_html(4)
make_alert_button(message =  "This is a test \\(X\\)", type = "hint2")
small_v_space_html(4)
make_alert_button(message = "Test \\( f_X(x) = \\frac{1}{\\sqrt{2 \\pi \\sigma^2}} \\exp\\Big(-\\frac{(x-\\mu)^2}{2\\sigma^2}\\Big) \\)", type = "explanation")
```

