---
title: "Linear regression: Exercises"
format: 
  live-html:
    theme: cosmo
    css: ../assets/webex_2.css
    include-after-body: ../assets/webex_2.js
engine: knitr
---

{{< include ../_extensions/r-wasm/live/_knitr.qmd >}}








```{r, echo = FALSE, results='asis'}
knitr::opts_chunk$set(echo = FALSE)
source("../assets/q_a_env.R")
```

<!-- MathJax for Rendering Math -->
<script type="text/javascript" async
  src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/3.2.2/es5/tex-mml-chtml.js">
</script>


### Question 1
Suppose that you want to compare the means of $2$ groups, in which situation is Student's t-test a good approach:


```{r, results='asis'}
# Define questions choice and individual feedback for each question
opts <- list(
  list(value = "We need the following: (1) it is reasonable to assume that there are no outliers and (2) for moderate and small sample sizes the sample distribution should be at least approximately normal.",
       feedback = "Not enough! We also need that the variances are the same between groups."),
  list(value = "We need the following: (1) it is reasonable to assume that there are no outliers and (2) for moderate and small sample sizes the sample distribution should be at least approximately normal. If these criteria are met then we should test if the variances are the same between groups. If we can't reject the null, then we can use Student's t-test.",
       feedback = "Nope, even though we can't reject the null, we are still not sure if the null is correct 😉. "),
  list(
    value = "We should do the following tests: (1) test if there are no outliers, (2) test if the data are normally distributed (one test per sample) and (3) test if the variances are the same. If we can't reject the null for all tests, then we should use Student's t-test.",
    feedback = "Nope, even though we can't reject the null, we are still not sure if the null is correct 😉."
  ),
  answer = list(
    value = "We need the following: (1) it is reasonable to assume that there are no outliers, (2) the variances of the two groups are known to be the same (or at least it is reasonable to assume that this is the case) and (3) for moderate and small sample sizes the sample distribution should be at least approximately normal.",
    feedback = "Yes! 😎"
  )
)
# render QA environment
render_q_a_env(opts)
```

### Question 1
How to plot the data?

Complete the code below to obtain the boxplot to compare the empirical distributions of the weight loss of the two diets:


:::: {.panel-tabset}

## Exercise



```{pyodide}
#| exercise: ex_1
import pandas as pd
import matplotlib.pyplot as plt

diet = pd.read_csv("https://raw.githubusercontent.com/ELSTE-Master/Data-Science/main/Data/diet.csv")
# compute weight loss
diet["weight.loss"] = diet["initial.weight"] - diet["final.weight"]

# Variable of interest
dietA = diet["weight.loss"][diet["diet.type"]=="A"]
dietC = diet["weight.loss"][diet["diet.type"]=="C"]

# Side-by-side boxplot
plt.figure(figsize=(6,4))
plt.______([dietA, dietC],
            labels=["Diet A", "Diet C"],
            patch_artist=True,  # color the boxes
            boxprops=dict(facecolor="lightblue", color="blue"),
            medianprops=dict(color="red"))

plt.ylabel("Weight loss (kg)")
plt.title("Comparison of Weight Loss for Diets A and C")
plt.show()
```

## Hints

::: { .hint exercise="ex_1"}
Consider using the `boxplot()` function from `matplotlib.pyplot`.

:::

## Solution

::: { .solution exercise="ex_1" }
Use  the `boxplot()` function from `matplotlib.pyplot`:


```python
import pandas as pd
import matplotlib.pyplot as plt

diet = pd.read_csv("https://raw.githubusercontent.com/ELSTE-Master/Data-Science/main/Data/diet.csv")

# compute weight loss
diet["weight.loss"] = diet["initial.weight"] - diet["final.weight"]

# Variable of interest
dietA = diet["weight.loss"][diet["diet.type"]=="A"]
dietC = diet["weight.loss"][diet["diet.type"]=="C"]

# Side-by-side boxplot
plt.figure(figsize=(6,4))
plt.boxplot([dietA, dietC],    #<1>
            labels=["Diet A", "Diet C"],
            patch_artist=True,  # color the boxes
            boxprops=dict(facecolor="lightblue", color="blue"),
            medianprops=dict(color="red"))

plt.ylabel("Weight loss (kg)")
plt.title("Comparison of Weight Loss for Diets A and C")
plt.show()

```
1. Use `plt.boxplot()`
:::

::::
